Bow shocks, bow waves, and dust waves. III. Diagnostics

### 2019-06-22 papers/bs-bw-dw-03-submit/latexdiff/ms03-v2-expand.tex
Do latexdiff between the original and final versions of paper 3

\newcommand\Qp{\ensuremath{Q_{\text{p}}}}
\newcommand\Qpbar{\ensuremath{\bar{Q}_{\text{p}}}}
\newcommand{\grain}{\ensuremath{_{\text{d}}}}
\newcommand{\B}{\ensuremath{_{\scriptscriptstyle\text{B}}}}
\newcommand{\alfven}{\ensuremath{_{\scriptscriptstyle\text{A}}}}
\newcommand{\xsec}{\ensuremath{\sigma\grain}}
\newcommand\frad{\ensuremath{f_{\text{rad}}}}
\newcommand\fmax{\ensuremath{f_{\text{max}}}}
\newcommand\thm{\ensuremath{\theta_{\text{m}}}}
\newcommand\drag{\ensuremath{_{\text{drag}}}}
\newcommand{\gas}{\ensuremath{_{\text{gas}}}}
\newcommand{\wind}{\ensuremath{_{\text{w}}}}
\newcommand{\trap}{\ensuremath{_{\text{abs}}}}
\newcommand{\ke}{\ensuremath{_{\text{kin}}}}
\newcommand{\drift}{\ensuremath{_{\text{drift}}}}
\newcommand\rad{\ensuremath{_{\text{rad}}}}
\newcommand\Lya{\ensuremath{_{\text{Ly}\alpha}}}
\newcommand\Rmin{\ensuremath{R_{\scriptscriptstyle\text{min}}}}
% Why do I need both of these?
\newcommand\sound{\ensuremath{c_{\text{s}}}}
\newcommand\soundspeed{\ensuremath{c_{\text{s,gas}}}}
\newcommand\starstar{\ensuremath{_{**}}}
\newcommand\mmp{\ensuremath{_{\text{\tiny MMP83}}}}
\newcommand\Hab{\ensuremath{_{\text{\tiny Habing}}}}
\newcommand\IR{\ensuremath{_{\text{IR}}}}
\newcommand{\thD}{\(\theta^1\)\,Ori~D}
\newcommand\alphaB{\ensuremath{\alpha_{\text{B}}}}
\newcommand\shell{\ensuremath{_{\text{sh}}}}
\newcommand\M{\ensuremath{\mathcal{M}}}
\newcommand\hii{\ion{H}{ii}}

@@

In Figure~\ref{fig:All-sources-eta-tau} we show the resultant
diagnostic diagram: \(\eta\shell\) versus \(\tau\).  The horizontal axis
shows the fraction of the stellar radiative \emph{energy} that is
reprocessed by the bow shell, while the vertical axis shows the
fraction of stellar radiative \emph{momentum} that is imparted to the
shell, either directly by absorption, or indirectly by the stellar
wind (which is itself radiatively driven).  Radiatively supported bows
(DW, RBW, or RBS, or cases) should lie on the diagonal line
\(\eta\shell = ( Q_P / Q_{\text{abs}}) \tau \approx 1.25 \tau\), where we have used
the ratio of grain radiation pressure efficiency to absorption
efficiency found in the FUV band for the dust mixture shown in
Paper~II's Figure~6.  Wind-supported bows should lie above this line
and no bows should lie below the \(\eta\shell = \tau\) line, since
\(Q_P\) cannot be smaller than \(Q_{\text{abs}}\).

@@

@@

@@

\bibitem[\protect\citeauthoryear{{Henney} \& {Arthur}}{{Henney} \&
  {Arthur}}{2019a}]{Henney:2019a}
{Henney} W.~J.,  {Arthur} S.~J.,  2019a, arXiv e-prints, \href
  {http://adsabs.harvard.edu/abs/2019arXiv190303737H} {1903.03737 MNRAS
  submitted (Paper I)}

\bibitem[\protect\citeauthoryear{{Henney} \& {Arthur}}{{Henney} \&
  {Arthur}}{2019b}]{Henney:2019b}
{Henney} W.~J.,  {Arthur} S.~J.,  2019b, arXiv e-prints, \href
  {http://adsabs.harvard.edu/abs/2019arXiv190307774H} {1903.07774 MNRAS
  submitted (Paper II)}

@@

\bibitem[\protect\citeauthoryear{{Hoang}, {Tram}, {Lee}  \& {Ahn}}{{Hoang}
  et~al.}{2018}]{Hoang:2018a}
{Hoang} T.,  {Tram} L.~N.,  {Lee} H.,   {Ahn} S.-H.,  2018, arXiv, \href
  {http://adsabs.harvard.edu/abs/2018arXiv181005557H} {}

@@

@@

\bibitem[\protect\citeauthoryear{{Tarango-Yong} \& {Henney}}{{Tarango-Yong} \&
  {Henney}}{2018}]{Tarango-Yong:2018a}
{Tarango-Yong} J.~A.,  {Henney} W.~J.,  2018, \mnras, 477, 2431

@@

\bibitem[\protect\citeauthoryear{{Weingartner} \& {Draine}}{{Weingartner} \&
  {Draine}}{2001}]{Weingartner:2001a}
{Weingartner} J.~C.,  {Draine} B.~T.,  2001, \apj, 548, 296

@@

@@

\appendix
% start input ./app-k17-k18.tex
\section{Covariance of derived quantities from propagation of observational uncertainties}
\label{sec:comb-uncert-covar}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "dusty-bow-wave"
%%% End:
 % end input ./app-k17-k18.tex

>>>>

\newcommand\Qp{\ensuremath{Q_{\text{p}}}}
\newcommand\Qpbar{\ensuremath{\bar{Q}_{\text{p}}}}
\newcommand{\grain}{\ensuremath{_{\text{d}}}}
\newcommand{\B}{\ensuremath{_{\scriptscriptstyle\text{B}}}}
\newcommand{\alfven}{\ensuremath{_{\scriptscriptstyle\text{A}}}}
\newcommand{\xsec}{\ensuremath{\sigma\grain}}
\newcommand\frad{\ensuremath{f_{\text{rad}}}}
\newcommand\fmax{\ensuremath{f_{\text{max}}}}
\newcommand\thm{\ensuremath{\theta_{\text{m}}}}
\newcommand\drag{\ensuremath{_{\text{drag}}}}
\newcommand{\gas}{\ensuremath{_{\text{gas}}}}
\newcommand{\wind}{\ensuremath{_{\text{w}}}}
\newcommand{\trap}{\ensuremath{_{\text{abs}}}}
\newcommand{\ke}{\ensuremath{_{\text{kin}}}}
\newcommand{\drift}{\ensuremath{_{\text{drift}}}}
\newcommand\rad{\ensuremath{_{\text{rad}}}}
\newcommand\Lya{\ensuremath{_{\text{Ly}\alpha}}}
\newcommand\Rmin{\ensuremath{R_{\scriptscriptstyle\text{min}}}}
% Why do I need both of these?
\newcommand\sound{\ensuremath{c_{\text{s}}}}
\newcommand\soundspeed{\ensuremath{c_{\text{s,gas}}}}
\newcommand\starstar{\ensuremath{_{**}}}
\newcommand\mmp{\ensuremath{_{\text{\tiny MMP83}}}}
\newcommand\Hab{\ensuremath{_{\text{\tiny Habing}}}}
\newcommand\IR{\ensuremath{_{\text{IR}}}}
\newcommand{\thD}{\(\theta^1\)\,Ori~D}
\newcommand{\thC}{\(\theta^1\)\,Ori~C}
\newcommand\alphaB{\ensuremath{\alpha_{\text{B}}}}
\newcommand\shell{\ensuremath{_{\text{sh}}}}
\newcommand\M{\ensuremath{\mathcal{M}}}
\newcommand\hii{\ion{H}{ii}}

@@

In Figure~\ref{fig:All-sources-eta-tau} we show the resultant
diagnostic diagram: \(\eta\shell\) versus \(\tau\).  The horizontal axis
shows the fraction of the stellar radiative \emph{energy} that is
reprocessed by the bow shell, while the vertical axis shows the
fraction of stellar radiative \emph{momentum} that is imparted to the
shell, either directly by absorption, or indirectly by the stellar
wind (which is itself radiatively driven).  Radiatively supported bows
(DW, RBW, or RBS cases) should lie on the diagonal line
\(\eta\shell = ( Q_P / Q_{\text{abs}}) \tau \approx 1.25 \tau\), see discussion
following equation~\eqref{eq:tau-empirical}.  Wind-supported bows
should lie above this line and no bows should lie below the
\(\eta\shell = \tau\) line, since \(Q_P\) cannot be smaller than
\(Q_{\text{abs}}\).

@@

\bibitem[\protect\citeauthoryear{{Abel}, {Hoof}, {Shaw}, {Ferland}  \&
  {Elwert}}{{Abel} et~al.}{2008}]{Abel:2008a}
{Abel} N.~P.,  {Hoof} P.~A.~M.~v.,  {Shaw} G.,  {Ferland} G.~J.,   {Elwert} T.,
   2008, \apj, 686, 1125

@@

\bibitem[\protect\citeauthoryear{Bohren \& Huffman}{Bohren \&
  Huffman}{1983}]{Bohren:1983a}
Bohren C.~F.,  Huffman D.,  1983, Absorption and scattering of light by small
  particles.
Wiley-VCH

@@

\bibitem[\protect\citeauthoryear{{Henney} \& {Arthur}}{{Henney} \&
  {Arthur}}{2019a}]{Henney:2019a}
{Henney} W.~J.,  {Arthur} S.~J.,  2019a, \mnras, \href
  {http://adsabs.harvard.edu/abs/2019arXiv190303737H} {486, 3423 (Paper I)}

\bibitem[\protect\citeauthoryear{{Henney} \& {Arthur}}{{Henney} \&
  {Arthur}}{2019b}]{Henney:2019b}
{Henney} W.~J.,  {Arthur} S.~J.,  2019b, \mnras, \href
  {http://adsabs.harvard.edu/abs/2019arXiv190307774H} {486, 4423 (Paper II)}

@@

\bibitem[\protect\citeauthoryear{{Hoang}, {Tram}, {Lee}  \& {Ahn}}{{Hoang}
  et~al.}{2018}]{Hoang:2018a}
{Hoang} T.,  {Tram} L.~N.,  {Lee} H.,   {Ahn} S.-H.,  2018, arXiv, \href
  {http://adsabs.harvard.edu/abs/2018arXiv181005557H} {1810.05557}

@@

\bibitem[\protect\citeauthoryear{{Lanz} \& {Hubeny}}{{Lanz} \&
  {Hubeny}}{2003}]{Lanz:2003a}
{Lanz} T.,  {Hubeny} I.,  2003, \apjs, 146, 417

@@

\bibitem[\protect\citeauthoryear{{Tarango-Yong} \& {Henney}}{{Tarango-Yong} \&
  {Henney}}{2018}]{Tarango-Yong:2018a}
{Tarango-Yong} J.~A.,  {Henney} W.~J.,  2018, \mnras, 477, 2431 (Paper 0)

@@

\bibitem[\protect\citeauthoryear{{Weingartner} \& {Draine}}{{Weingartner} \&
  {Draine}}{2001a}]{Weingartner:2001b}
{Weingartner} J.~C.,  {Draine} B.~T.,  2001a, \apjs, 134, 263

\bibitem[\protect\citeauthoryear{{Weingartner} \& {Draine}}{{Weingartner} \&
  {Draine}}{2001b}]{Weingartner:2001a}
{Weingartner} J.~C.,  {Draine} B.~T.,  2001b, \apj, 548, 296

@@

\bibitem[\protect\citeauthoryear{{Wibking}, {Thompson}  \&
  {Krumholz}}{{Wibking} et~al.}{2018}]{Wibking:2018a}
{Wibking} B.~D.,  {Thompson} T.~A.,   {Krumholz} M.~R.,  2018, \mnras, 477,
  4665

@@

\appendix

\section{Dust opacity and optical depth}
\label{app:dust-opacity}

In the main body of the paper we use a frequency-averaged dust optical
depth, \(\tau\), which we here relate to the frequency-dependent optical
properties of the grains.  For simplicity, we assume a dust model that
consists of a set \(k \in \{1 \dots N\}\) of discrete populations of
solid spherical grains \citep[e.g.][]{Desert:1990a}, each with its own
chemical composition and size distribution:
\(y_{a,k} \equiv n_k^{-1} (d n_k / d a)\), where \(n_k\) is the volume
density of the \(k\)-th population (\si{grain.cm^{-3}}) and \(a\) is
the grain radius. The frequency-dependent dust opacity per unit mass
of gas and dust is \(\kappa_\nu\), with units \si{cm^2.g^{-1}}, which can be
expressed as
\begin{equation}
  \label{eq:dust-opacity-frequency}
  \kappa_\nu = \rho^{-1} \sum_{k=1}^{N} \left[  n_k \int_0^\infty \!\!y_{a,k} \, Q_{a,k,\nu} \,\pi a^2 \, da\right]\ ,
\end{equation}
where \(\rho\) is the total material mass density and \(Q_{a,k,\nu}\) is
the size-, composition-, and frequency-dependent dimensionless grain
optical efficiency, which can be calculated from the grain refractive
indices using Mie theory \citep{Bohren:1983a}.  There are different
varieties of \(Q\), and therefore different varieties of opacity,
according to whether one is interested in true absorption,
\(Q_{\text{abs}}\), scattering, \(Q_{\text{scat}}\), or total
extinction, \(Q_{\text{ext}} = Q_{\text{abs}} + Q_{\text{scat}}\).
The relative importance of scattering is described by the grain
albedo: \(\varpi = Q_{\text{scat}}/Q_{\text{ext}}\).  The radiation
pressure acting on the grains is determined by a fourth variety of
optical efficiency:
\(Q_P = Q_{\text{abs}} + (1 - g) Q_{\text{scat}}\), where \(g\) is the
mean scattering cosine.

The monochromatic optical depth of a medium (in any of these 4
varieties) along a path \(s\) is given by
\begin{equation}
  \label{eq:tau-monochromatic}
  \tau_\nu = \int \rho \kappa_\nu \, ds \ .
\end{equation}
If the effects of multiple scattering are ignored, then the fraction
of stellar radiation of frequency \(\nu\) that is absorbed by the shell
is \(1 - e^{-\tau_\nu}\), in which the integral is performed with
\(s\) along the radial direction, assuming the star to be effectively
a point source at the position of the shell.  To find the fraction of
radiative \emph{energy}, \(\kappa_\nu\) must be calculated using
\(Q_{\text{abs}}\), whereas to find the fraction of radiative
\emph{momentum}, it must be calculated using \(Q_P\).  This
single-scattering regime \citep[e.g.][]{Wibking:2018a} is a good
approximation so long as \(\varpi\) is not close to unity and the
shell is optically thin at infrared wavelengths, which is always true
for stellar bow shocks.

The frequency-averaged optical depth \(\tau\) is defined such that the
fraction of the frequency-integrated stellar radiative momentum
absorbed by the shell is \(1 - e^{-\tau}\).  For a stellar spectrum of
shape \(F_\nu\), this implies the formal relation
\begin{equation}
  \label{eq:tau-averaged}
  \tau = - \ln \left[ \frac{\int_0^\infty F_\nu \, e^{-\tau_\nu} \, d\nu}{\int_0^\infty F_\nu  \, d\nu} \right] \ .
\end{equation}
Note that this equation is valid regardless of the physical units of
\(F_\nu\) (e.g., flux or luminosity).  More approximately, \(\tau\) can
simply be estimated by \(\tau_\nu\) at the peak frequency of the stellar
radiation, which typically falls in the far-ultraviolet range,
although this approximation breaks down if \(\tau_\nu > 1\) over a
significant portion of the stellar spectrum.

In the main body of the paper, we use \(\tau_P\) for the radiation pressure
optical depth and \(\tau_{\text{abs}}\) for the absorption optical depth.  These are
related by the ratio of the corresponding frequency-averaged optical
efficiencies:
\begin{equation}
  \label{eq:tau_tauprime}
  \tau_P = \frac{Q_P}{Q_{\text{abs}}} \tau_{\text{abs}} = \left(  1 + \varpi \frac{1 - g}{1 - \varpi}\right) \tau_{\text{abs}}\ .
\end{equation}
Specific results were given in Figure~6a of Paper~II for an example
grain mixture, designed to reproduce average interstellar medium dust
properties \citep{Weingartner:2001b, Abel:2008a, Ferland:2013a,
  Ferland:2017a}.  For such a mixture the albedo is roughly constant
at \(\varpi \approx 0.5\) throughout the ultraviolet and visible bands,
falling to low values in the infrared.  The asymmetry parameter, or
mean scattering cosine, falls across the ultraviolet band from
\(g \approx 1\) at the highest frequencies to \(g\approx 0.5\) in the FUV.\@ This
yields a conversion factor of
\(Q_P / Q_{\text{abs}} = \text{\numrange{1.2}{1.3}}\) at EUV/FUV
wavelengths.

% start input ./app-k17-k18.tex
\section{Covariance of derived quantities from propagation of observational uncertainties}
\label{sec:comb-uncert-covar}

@@

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "bs-bw-dw-03"
%%% End:
 % end input ./app-k17-k18.tex
